Enforced Invariant Confluence

### 2018-03-03 doc/paper_redux/iconfluence.tex
Notational changes and writing improvements.

\subsection{State-Based}
\begin{definition}
  A \defword{distributed state-based object} is a triple $(S, s_0, \join)$
  where $S$ is a set of states, $s_0 \in S$ is a designated start state, and
  $\join: S \times S \to S$ is a binary merge operator that we call join.
\end{definition}

The notion of state-based objects is taken from~\cite{shapiro2011conflict}.
Note that $\join$ is an \emph{arbitrary} function. It does not necessarily have
to satisfy any special properties like associativity or commutativity, though
later we will see that often it does.

@@

\begin{example}
  $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
  $\setst{x \in \nats}{x \equiv 0 \mod 2}$ is an invariant. Here, $I(0)$ and
  $I(2)$ but $\lnot I(1)$ and $\lnot I(3)$.
\end{example}

There are three ways to think about state-based \Iconfluence{}: a process-based
approach, a graph-based approach, and an expression-based approach. All three
approaches are illustrated in \figref{StateBasedModels} and are described in
more detail below.

@@

\subsubsection{Process-Based}
The process-based model is most similar to the process model described by
Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a set
$\seq{p}{1}{n}$ of $n$ processors each of which begins with state $s_0$.  A
processor $p_i$ can perform one of two actions.

First, $p_i$ can execute a transaction $t \in T$ to transition from state $s$
to state $t(s)$, given that $I(t(s))$. If $\lnot I(t(s))$, then $p_i$ will not
execute $t$ (or it will execute $t$ but then abort it; think of it however you
like).

Second, $p_i$ can send its state $s_i$ to another processor $p_j$ with state
$s_j$ causing $p_j$ to transition from state $s_j$ to state $s_i \join s_j$ (or
$s_j \join s_i$; it doesn't really matter). Note that unlike executing a
transaction, $p_j$ must transition from $s_j$ to $s_i \join s_j$ even if $\lnot
I(s_i \join s_j)$.

$T$ is invariant-confluent with respect to $I$, abbreviated \Iconfluent{}, if
every reachable state (including $s_0$) satisfies the invariant. Note that this
definition of \Iconfluence{} is different from the original definition
in~\cite{bailis2014coordination}. Later, we will see that they are equivalent.

\subsubsection{Graph-Based}
The graph-based approach is most similar to the model used by Bailis et al.\
in~\cite{bailis2014coordination}. We are given a directed acyclic graph where
vertices are states, and edges are either labelled with transactions or are
unlabelled if they correspond to a join. The graph is really just an alternate
way of representing an execution in the process-based model, but with duplicate
states collapsed into a single vertex. As with the process-based model, we say
that $T$ is \Iconfluent{} if every vertex in every graph satisfies the
invariant.

\subsubsection{Expression-Based}
The expression-based approach formalizes the process-based and graph-based
approach using expressions. Given a state based object $(S, s_0, \join)$ and a
set of transactions $T$, we consider expressions generated by the following
grammar:
\[
  e ::= s_0 \mid t(e) \mid e_1 \join e_2
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression
$e$, written $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

@@

\begin{definition}
  $T$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq I$. In
  other words, $T$ is \Iconfluent{} if all reachable states satisfy the
  invariant.
\end{definition}

@@

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \Irec{s_0} \defeq I(s_0)

    \Irec{t(e)} \defeq I(t(e)) \land \Irec{e}

    \Irec{e_1 \join e_2}
    \defeq I(e_1 \join e_2) \land \Irec{e_1} \land \Irec{e_2}
  \end{mathpar}
\end{definition}

Bailis et al.\ defined \Iconfluence{} to mean that expressions recursively
satisfying $I$ are closed under join. That is, $T$ is \Iconfluent{} if
\[
  \forall e_1, e_2.\; \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)
\]
If $\lnot I(s_0)$, then \Iconfluence{} holds vacuously which is a bit silly, so
Bailis et al.\ probably ought to have also added the condition $I(s_0)$ to
their definition of \Iconfluence{}. Doing so, the two definitions become
equivalent.

\begin{claim}\clmlabel{StateBasedIrecImpliesReachable}
  $\Irec{e} \implies \reachable{e}$
\end{claim}
\begin{elidableproof}
  We perform structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

@@

\begin{claim}\clmlabel{StateBasedTwoIconfluenceDefs}
  Consider a state based object $(S, s_0, \join)$, a set of transactions $T$,
  and an invariant $I$. The following two are equivalent:
  \begin{enumerate}[\quad(1)]
    \item
      $\setst{e}{\reachable{e}} \subseteq I$

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)$
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). Axiomatically, $\reachable{s_0}$, so by
  (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{StateBasedIrecImpliesReachable},
  $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
  so by (1), $I(e_1 \join e_2)$.

@@

\subsection{Operation-Based}
\begin{definition}
  A \defword{distributed operation-based object} is a pair $(S, s_0)$ where $S$
  is a set of states and $s_0 \in S$ is a designated start state.
\end{definition}

@@

\begin{example}
  $(\nats, 0)$ is a distributed object. $t(x) = y \mapsto x + y$ is a
  transaction that given a state $x$ returns a function $y \mapsto x + y$ that
  adds $x$ to its argument. $\setst{x \geq 0}{x \in \nats}$ is an invariant.
\end{example}

@@

\subsubsection{Process-Based}
As with the state-based approach, the process-based model is most similar to
the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
state $s_0$. A processor $p_i$ can do one of two things.

First, $p_i$ can execute a transaction $t$ on its current state $s$ and then
transitions from state $s$ to state $t(s)(s)$, given that $I(t(s)(s))$. If
$\lnot I(t(s)(s))$, then $p_i$ will not perform $t$. If $p_i$ does execute $t$,
then it broadcasts $t(s)$ to all other processors exactly once.

Second, $p_i$ can receive a broadcasted shadow transaction $t(s_j)$ from some
other processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
state $s_i$ to state $t(s_j)(s_i)$. When $p_i$ receives a shadow transaction,
it must execute it, even if $\lnot I(t(s_j)(s_i))$.

$T$ is \Iconfluent{} if every reachable state (including $s_0$) satisfies the
invariant.

\subsubsection{Graph-Based}
In the graph-based model, we are given a directed acyclic graph in which each
vertex is a state $s$ and each edge is labelled with a shadow operation $t(s)$
where $s$ is some other vertex in the graph. $T$ is \Iconfluent{} if every
vertex in every graph satisfies $I$.

\subsubsection{Expression-Based}
Given an operation-based object $(S, s_0)$ and a set of transactions $T$, we
consider expressions built from the following grammar:
\[
  e ::= s_0 \mid t(e_1)(e_2)
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
denoted $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

@@

\begin{definition}
  $T$ is \Iconfluent{} if $\setst{e}{\reachable{e}} \subseteq I$.
\end{definition}

@@

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \Irec{s_0} \defeq I(s_0)

    \Irec{t(e_1)(e_2)} \defeq I(t(e_1)(e_2)) \land \Irec{e_1} \land \Irec{e_2}
  \end{mathpar}
\end{definition}

\begin{claim}\clmlabel{OpBasedIrecImpliesReachable}
  $\Irec{e} \implies \reachable{e}$
\end{claim}
\begin{elidableproof}
  We perform a structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

@@

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \land I(t(e_1)(e_1)) \implies
      I(t(e_1)(e_2))$.
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). $\reachable{s_0}$, so by (1), $I(s_0)$.
  Let $e_1$ and $e_2$ be arbitrary expressions such that $\Irec{e_1}$,
  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \clmref{OpBasedIrecImpliesReachable},
  $\reachable{e_1}$ and $\reachable{e_2}$, so $\reachable{t(e_1)(e_2)}$. By
  (1), $I(t(e_1)(e_2))$.

>>>>

\subsection{State-Based}
\begin{definition}
  A \defword{distributed state-based object} $O$ is a triple $(S, s_0, \join)$
  where $S$ is a set of states, $s_0 \in S$ is a designated start state, and
  $\join: S \times S \to S$ is a binary merge operator that we call join.
\end{definition}

The notion of state-based objects is taken from~\cite{shapiro2011conflict}.
Note that $\join$ is an \emph{arbitrary} function. It does not necessarily have
to satisfy any special properties like associativity or commutativity.

@@

We formalize an invariant as a set of states, but we can also think of a
predicate as a formula over states. For example, we can think of the invariant
$\setst{(x, y) \in \ints \times \ints}{x \geq y}$ as the formula $x \geq y$ and
vice-versa.

\begin{example}
  $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
  $\setst{x \in \nats}{\text{$x$ is even}}$ is an invariant. Here, $I(0)$ and
  $I(2)$ but $\lnot I(1)$ and $\lnot I(3)$.
\end{example}

There are three ways to think about state-based \Iconfluence{}: a process-based
approach, a graph-based approach, and an expression-based approach. Below, we
describe the process-based approach and the graph-based approach informally and
describe the expression-based approach formally. All three approaches are
illustrated in \figref{StateBasedModels}.

@@

\subsubsection{Process-Based}
The process-based model is most similar to the process model described by
Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a state-based
object $O = (S, s_0, \join)$, a set of transactions $T$, an invariant $I$, and
a set $\seq{p}{1}{n}$ of $n$ processors. Each processor begins with state
$s_0$. A processor $p_i$ can perform one of two actions.

\begin{itemize}
  \item
    $p_i$ can execute a transaction $t \in T$ to transition from state $s$ to
    state $t(s)$, given that $I(t(s))$. If $\lnot I(t(s))$, then $p_i$ will not
    execute $t$ (or it will execute $t$ but then abort it; think of it however
    you like).

  \item
    $p_i$ can send its state $s_i$ to another processor $p_j$ with state $s_j$
    causing $p_j$ to transition from state $s_j$ to state $s_i \join s_j$ (or
    $s_j \join s_i$; it doesn't really matter). Note that unlike executing a
    transaction, $p_j$ must transition from $s_j$ to $s_i \join s_j$ even if
    $\lnot I(s_i \join s_j)$.
\end{itemize}

$(O, T)$ is invariant-confluent with respect to $I$ (abbreviated \Iconfluent{})
if every reachable state (including $s_0$) satisfies the invariant. Note that
this definition of \Iconfluence{} is different from the original definition
in~\cite{bailis2014coordination}. Later, we will see that they are equivalent.

\subsubsection{Graph-Based}
The graph-based approach is most similar to the model used by Bailis et al.\
in~\cite{bailis2014coordination}. We are given a directed acyclic graph where
vertices are states, and edges are either labelled with transactions or are
unlabelled if they correspond to a join. The graph is really just an alternate
way of representing an execution in the process-based model, but with duplicate
states collapsed into a single vertex. As with the process-based model, we say
that $(O, T)$ is \Iconfluent{} if every vertex in every graph satisfies the
invariant.

\subsubsection{Expression-Based}
The expression-based approach formalizes the process-based and graph-based
approach using expressions. Given a state based object $O = (S, s_0, \join)$
and a set of transactions $T$, we consider expressions generated by the
following grammar:
\[
  e ::= s_0 \mid t(e) \mid e_1 \join e_2
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression
$e$, written $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

@@

\begin{definition}
  $(O, T)$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq
  I$. In other words, $T$ is \Iconfluent{} if all reachable states satisfy the
  invariant.
\end{definition}

@@

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \inferrule{I(s_0)}{\Irec{s_0}}

    \inferrule{I(t(e)) \\ \Irec{e}}{\Irec{t(e)}}

    \inferrule{I(e_1 \join e_2) \\ \Irec{e_1} \\ \Irec{e_2}}
              {\Irec{e_1 \join e_2}}
  \end{mathpar}
\end{definition}

Bailis et al.\ defined \Iconfluence{} to mean that expressions recursively
satisfying $I$ are closed under join. That is, $(O, T)$ is \Iconfluent{} if
\[
  \forall e_1, e_2.\; \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)
\]
If $\lnot I(s_0)$, then \Iconfluence{} holds vacuously which is a bit silly, so
Bailis et al.\ probably ought to have also added the condition $I(s_0)$ to
their definition of \Iconfluence{}. Doing so, the two definitions become
equivalent.

\begin{lemma}\lemlabel{StateBasedIrecImpliesReachable}
  $\Irec{e} \implies \reachable{e}$
\end{lemma}
\begin{elidableproof}
  We perform structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

@@

\begin{claim}\clmlabel{StateBasedTwoIconfluenceDefs}
  Consider a state based object $O = (S, s_0, \join)$, a set of transactions
  $T$, and an invariant $I$. The following two are equivalent:
  \begin{enumerate}[\quad(1)]
    \item
      $\setst{e}{\reachable{e}} \subseteq I$

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)$
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). Axiomatically, $\reachable{s_0}$, so by
  (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
  $\Irec{e_1}$ and $\Irec{e_2}$. By \lemref{StateBasedIrecImpliesReachable},
  $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
  so by (1), $I(e_1 \join e_2)$.

@@

\subsection{Operation-Based}
\begin{definition}
  A \defword{distributed operation-based object} is a pair $O = (S, s_0)$ where
  $S$ is a set of states and $s_0 \in S$ is a designated start state.
\end{definition}

@@

\begin{example}
  $(\nats, 0)$ is a distributed object. $t(x) = \lambda y.\ x + y$ is a
  transaction that given a state $x$ returns a function $\lambda y.\ x + y$
  that adds $x$ to its argument. $\setst{x \in \nats}{x \geq 0}$ is an
  invariant.
\end{example}

@@

\subsubsection{Process-Based}
As with the state-based approach, the process-based model is most similar to
the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
We are given an operation-based object $O$, a set of transactions $T$, an
invariant $I$, and a set $\seq{p}{1}{n}$ of $n$ processors. Each processors
begins with state $s_0$. A processor $p_i$ can do one of two things.

\begin{itemize}
  \item
    $p_i$ can execute a transaction $t$ on its current state $s$ and then
    transitions from state $s$ to state $t(s)(s)$, given that $I(t(s)(s))$. If
    $\lnot I(t(s)(s))$, then $p_i$ will not perform $t$. If $p_i$ does execute
    $t$, then it broadcasts $t(s)$ to all other processors exactly once.

  \item
    $p_i$ can receive a broadcasted shadow transaction $t(s_j)$ from some other
    processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
    state $s_i$ to state $t(s_j)(s_i)$. When $p_i$ receives a shadow
    transaction, it must execute it, even if $\lnot I(t(s_j)(s_i))$.
\end{itemize}

$(O, T)$ is \Iconfluent{} if every reachable state (including $s_0$) satisfies
the invariant.

\subsubsection{Graph-Based}
In the graph-based model, we are given a directed acyclic graph in which each
vertex is a state $s$ and each edge is labelled with a shadow operation $t(s)$
where $s$ is some other vertex in the graph. $(O, T)$ is \Iconfluent{} if every
vertex in every graph satisfies $I$.

\subsubsection{Expression-Based}
Given an operation-based object $O = (S, s_0)$ and a set of transactions $T$,
we consider expressions built from the following grammar:
\[
  e ::= s_0 \mid t(e_1)(e_2)
\]
where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
denoted $eval(e)$, in the obvious way:
\begin{mathpar}
  eval(s_0) \defeq s_0

@@

\begin{definition}
  $(O, T)$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq
  I$.
\end{definition}

@@

\begin{definition}
  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
  and all of $e$'s children satisfy $I$. That is,
  \begin{mathpar}
    \inferrule{I(s_0)}{\Irec{s_0}}

    \inferrule{I(t(e_1)(e_2)) \\ \Irec{e_1} \\ \Irec{e_2}}{\Irec{t(e_1)(e_2)}}
  \end{mathpar}
\end{definition}

\begin{lemma}\lemlabel{OpBasedIrecImpliesReachable}
  $\Irec{e} \implies \reachable{e}$
\end{lemma}
\begin{elidableproof}
  We perform a structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case $s_0$.}
      Axiomatically, $\reachable{s_0}$.

@@

    \item
      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \land I(t(e_1)(e_1)) \implies
      I(t(e_1)(e_2))$.
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  First, we show that (1) implies (2). $\reachable{s_0}$, so by (1), $I(s_0)$.
  Let $e_1$ and $e_2$ be arbitrary expressions such that $\Irec{e_1}$,
  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \lemref{OpBasedIrecImpliesReachable},
  $\reachable{e_1}$ and $\reachable{e_2}$, so $\reachable{t(e_1)(e_2)}$. By
  (1), $I(t(e_1)(e_2))$.

### 2018-03-03 doc/paper_redux/interactive_decision_procedure.tex
Notational changes and writing improvements.

\section{Interactive \Iconfluence{} Decision Procedure}
Imagine we have an \Iclosed{} decision procedure, and let's revisit
\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a set of
transactions $T$, and an invariant $I$. We want to know if $T$ is
\Iconfluent{}. We can hand $O$, $T$, and $I$ to the \Iclosed{} decision
procedure to see if $O$ is \Iclosed{}. If the decision procedure says yes, then
by \clmref{StateBasedSufficient}, we know that $T$ is \Iconfluent{}.

However, in our example, $O$ is not \Iclosed{}, so the decision procedure will
say no. \clmref{SufficientAndNecessary} tells us that this means one of two
things. Either (1) $T$ is not \Iconfluent{} or (2) $T$ is \Iconfluent{}, but
there are some invariant satisfying states that are not reachable. If we think
harder about the set of reachable states, we may realize that all reachable
states have a non-negative value of $x$. Thus, we may modify our invariant to
$\setst{(x, y)}{xy \leq 0 \land x \geq 0}$. Now, $O$ is \Iclosed{}! If we hand
$O$, $T$, and $I$ to the decision procedure, it will say yes, and we can
conclude that $T$ is \Iconfluent{}.

\newcommand{\comment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}
  \caption{Interactive \Iconfluence{} Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsIconfluent}{$O$, $T$, $I$}
      \State \Return \Call{Helper}{$O$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

@@

    \comment{$R$ is a set of reachable states.}
    \comment{$NR$ is a set of non-reachable states.}
    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
      \State closed, $s_1$, $s_2$ $\gets$
               \Call{IsIclosed}{$O$, $T$, $I - NR$}
      \If {closed $=$ unknown}
        \State \Return unknown
      \ElsIf {closed $=$ yes}
        \State \Return{yes}
      \Else
        \State Augment $R$ and $NR$ with random search and user input
        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
        \If{$s_1, s_2 \in R$}
          \State \Return no
        \Else
          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
        \EndIf
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

This example motivates \algoref{InteractiveDecisionProcedure}, a general
purpose interactive \Iconfluence{} decision procedure in which a user
iteratively refines the invariant until the decision procedure terminates.
Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
\IsIconfluent$(O, T, I)$ to determine if $T$ is \Iconfluent{}.

\IsIconfluent{} assumes access to an \Iclosed{} decision procedure
\IsIclosed$(O, T, I)$ which returns whether $O$ is \Iclosed{} (yes, no, or
unknown). If it returns no, it also returns a counterexample $s_1$ and $s_2$
such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$.  Constructing an
\Iclosed{} decision procedure is hard (undecidable, I think), but we can do
pretty well using an SMT solver like Z3. There is also some existing research
on how to built smart \Iclosed{} decision procedures that we can
leverage~\cite{li2014automating}.

@@

    After $s_1$ and $s_2$ have been labelled reachable or not, we continue. If
    both $s_1$ and $s_2$ are reachable, then so is $s_1 \join s_2$, but $\lnot
    I(s_1 \join s_2)$. Thus, $T$ is not \Iconfluent{}, so \Helper{} returns no.
    Otherwise, one of $s_1$ and $s_2$ is unreachable. Now, we recurse.
\end{itemize}

Let's walk through how we might use \IsIconfluent{} on
\exampleref{StateBasedNotNecessary}.
\begin{itemize}
  \item
    \IsIconfluent{} calls \Helper{} with $R = {(0, 0)}, NR = \emptyset{}$.

@@

>>>>

\section{Interactive \Iconfluence{} Decision Procedure}
Imagine we have an \Iclosed{} decision procedure, and let's revisit
\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a set of
transactions $T$, and an invariant $I$. We want to know if $T$ is
\Iconfluent{}. Assume we have already determined that $I(s_0)$. We can hand $O$
and $I$ to the \Iclosed{} decision procedure to see if $O$ is \Iclosed{}. If
the decision procedure says yes, then by \clmref{StateBasedSufficient}, we know
that $(O, T)$ is \Iconfluent{}.

However, in our example, $O$ is not \Iclosed{}, so the decision procedure will
say no. \clmref{SufficientAndNecessary} tells us that this means one of two
things. Either (1) $T$ is not \Iconfluent{} or (2) $(O, T)$ is \Iconfluent{},
but there are some invariant satisfying states that are not reachable. If we
think harder about the set of reachable states, we may realize that all
reachable states have a non-negative value of $x$. Thus, we may modify our
invariant from $\setst{(x, y)}{xy \leq 0}$ to $\setst{(x, y)}{xy \leq 0 \land x
\geq 0}$. Now, $O$ is \Iclosed{}! If we hand $O$ and $I$ to the decision
procedure, it will say yes, and we can conclude that $(O, T)$ is \Iconfluent{}.

\newcommand{\comment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}
  \caption{Interactive \Iconfluence{} Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsIconfluent}{$O$, $T$, $I$}
      \If {$\lnot I(s_0)$}
        \State \Return no
      \Else
        \State \Return \Call{Helper}{$O$, $T$, $I$, $\set{s_0}$, $\emptyset$}
      \EndIf
    \EndFunction

@@

    \comment{$R$ is a set of reachable states.}
    \comment{$NR$ is a set of non-reachable states.}
    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
      \If {closed $=$ unknown}
        \State \Return unknown
      \ElsIf {closed $=$ yes}
        \State \Return{yes}
      \Else
        \State Augment $R$ and $NR$ with random search and user input
        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
        \If{$s_1, s_2 \in R$}
          \State \Return no
        \Else
          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
        \EndIf
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

This example motivates \algoref{InteractiveDecisionProcedure}, a general
purpose interactive \Iconfluence{} decision procedure in which a user
iteratively refines the invariant until the decision procedure terminates.
Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
\IsIconfluent$(O, T, I)$ to determine if $(O, T)$ is \Iconfluent{}.

\IsIconfluent{} assumes access to an \Iclosed{} decision procedure
\IsIclosed$(O, I)$ which returns whether $O$ is \Iclosed{} (yes, no, or
unknown). If it returns no, it also returns a counterexample $s_1$ and $s_2$
such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$. Constructing an
\Iclosed{} decision procedure is hard (undecidable, I think), but we can do
pretty well using an SMT solver like Z3. There is also some existing research
on how to built smart \Iclosed{} decision procedures that we can
leverage~\cite{li2014automating}.

@@

    After $s_1$ and $s_2$ have been labelled reachable or not, we continue. If
    both $s_1$ and $s_2$ are reachable, then so is $s_1 \join s_2$, but $\lnot
    I(s_1 \join s_2)$. Thus, $(O, T)$ is not \Iconfluent{}, so \Helper{}
    returns no.  Otherwise, one of $s_1$ and $s_2$ is unreachable. Now, we
    recurse.
\end{itemize}

Let's walk through how we might use \IsIconfluent{} on
\exampleref{StateBasedNotNecessary}.
\begin{itemize}
  \item
    \IsIconfluent{} calls \Helper{} with $R = \set{(0, 0)}, NR = \emptyset{}$.

@@

\todo{What if the user labels things incorrectly?}

### 2018-03-03 doc/paper_redux/new_statebased_sufficient.tex
Made some proofs elidable.
Notational changes and writing improvements.

\section{New Sufficient State-Based Conditions}
\todo{Motivate these in terms of the interactive decision procedure.}

In this section, we outline a new sufficient condition for state-based
\Iconfluence{} that is more general than the one in
\clmref{StateBasedSufficient}. To prove that the condition is sufficient, we
introduce the concepts of \dIconfluence{} and \dIreduction{}.

@@

\begin{claim}\clmlabel{DiconfluenceDireducibleImpliesIconfluence}
  If $T$ is \dIconfluent{} and \dIreducible, then it is \Iconfluent.
\end{claim}
\begin{proof}
  First, we prove that for all $e$, if $\Irec{e}$ then there exists a join-free
  $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a pretty
  straightforward structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case 1: $e = s_0$.}
      Trivially, $e' = s_0.$

@@

  Next, consider arbitrary expressions $e_1$ and $e_2$ where $\Irec{e_1}$ and
  $\Irec{e_2}$. To show that $T$ is \Iconfluent, we must show that $I(e_1 \join
  e_2)$. Using our little lemma above, we know there exists join-free $e_1'$
  and $e_2'$ where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') =
  eval(e_2)$, and $\Irec{e_2'}$. Because $T$ is \dIconfluent, $I(e_1' \join
  e_2')$, and
  \[
    eval(e_1' \join e_2')
      = eval(e_1') \join eval(e_2')
      = eval(e_1) \join eval(e_2)
      = eval(e_1 \join e_2)
  \]
  Thus, $I(e_1 \join e_2)$.
\end{proof}

@@

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.
  \end{enumerate}
\end{claim}
\begin{proof}
  \newcommand{\bart}[1]{\overline{t_{#1}}}
  \newcommand{\baru}[1]{\overline{u_{#1}}}

@@

  Next, we show that $T$ is \dIreducible{}. Again let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free expressions that recursively satisfy $I$. We just
  proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also, $eval(e_1 \join e_2) =
  eval(\bart{n}(\baru{m}(s_0)))$. $\bart{n}(\baru{m}(s_0))$ is a join-free
  expression, so $T$ is \dIreducible{}.
\end{proof}

>>>>

\section{New Sufficient State-Based Conditions}
\todo{%
  Revise everything with new notation of $(O, T)$ \Iconfluent{} instead of $T$
  \Iconfluent{}.
}

\algoref{InteractiveDecisionProcedure} relies on a user to provide information
about reachable and unreachable states. In this section, we outline a new
sufficient condition for state-based \Iconfluence{} that does not require user
interaction and that covers some cases not covered by the sufficient conditions
in \clmref{StateBasedSufficient}. To prove that the new condition is
sufficient, we introduce the concepts of \dIconfluence{} and \dIreduction{}.

@@

\begin{claim}\clmlabel{DiconfluenceDireducibleImpliesIconfluence}
  If $T$ is \dIconfluent{} and \dIreducible, then it is \Iconfluent.
\end{claim}
\begin{elidableproof}
  First, we prove that for all $e$, if $\Irec{e}$ then there exists a join-free
  $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a pretty
  straightforward structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case 1: $e = s_0$.}
      Trivially, $e' = s_0.$

@@

  Next, consider arbitrary expressions $e_1$ and $e_2$ where $\Irec{e_1}$ and
  $\Irec{e_2}$. To show that $T$ is \Iconfluent, we must show that $I(e_1 \join
  e_2)$. Using our little lemma above, we know there exists join-free $e_1'$
  and $e_2'$ where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') =
  eval(e_2)$, and $\Irec{e_2'}$. Because $T$ is \dIconfluent, $I(e_1' \join
  e_2')$, and
  \[
    eval(e_1' \join e_2')
      = eval(e_1') \join eval(e_2')
      = eval(e_1) \join eval(e_2)
      = eval(e_1 \join e_2)
  \]
  Thus, $I(e_1 \join e_2)$.
\end{elidableproof}

@@

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  \newcommand{\bart}[1]{\overline{t_{#1}}}
  \newcommand{\baru}[1]{\overline{u_{#1}}}

@@

  Next, we show that $T$ is \dIreducible{}. Again let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free expressions that recursively satisfy $I$. We just
  proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also, $eval(e_1 \join e_2) =
  eval(\bart{n}(\baru{m}(s_0)))$. $\bart{n}(\baru{m}(s_0))$ is a join-free
  expression, so $T$ is \dIreducible{}.
\end{elidableproof}
